---
title: "Project2"
author: "LeTicia Cancel"
date: "9/27/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
Below is a analysis of different data sets provided by classmates. This analysis will attempt to answer specific questions from the person who provided the data, or a question that seems appropriate during data tidying. 

Libraries needed for analysis.
```{r message=FALSE, warning=FALSE}
library(sf)
library(DBI)
library(RCurl)
library(readr)
library(ggplot2)
library(dplyr)
library(tidyr)
```

# Data 1 - NYC Restaurant Inspection 

### Get Dataset 

The first data set that will be analyzed are all of the NYC Restaurant Inspection results from 2019 to 10/1/2020. The data will be pulled from a csv file that is on GitHub.
```{r message=FALSE,warning=FALSE}
#get URL and import data from GitHub
site <- "https://raw.githubusercontent.com/ltcancel/Project2_Data607_F20/master/DOHMH_New_York_City_Restaurant_Inspection_Results.csv"
df <- read_csv(url(site))

head(df)
```
### Tidy the data/Analysis

Q. How many violations were issued for restaurants in each borough for each year? 
```{r}
#separate date into its own columns
df <- df %>%
  separate('INSPECTION DATE',sep="/",into = c("month","day","year"),remove = FALSE)

#not every row has a value for BORO so I want to exclude this data. I also want to include anyone who does not have a violation description
df1Filter <- df %>% group_by(year) %>% count(BORO) %>% filter(BORO != 0 & !is.na('VIOLATION DESCRIPTION'))

head(df1Filter)
```

After plotting the data, we can see that the number of violations has increased over the years. Manhattan consistently has the greatest number of violations. Brooklyn and Queens are very close, and Staten Island has the least number of violations. I think this data needs further analysis to see the percentage of restaurants with violations and the percentage of restaurants with an A grade. Manhattan may have a larger number of restaurants making their violations seem inflated when compared to the other boroughs. 

```{r}
ggplot(df1Filter) +
  geom_bar(mapping = aes(x=BORO, y = n, fill = BORO), stat = 'identity') +
  facet_wrap(~year) +
  coord_flip() +
  labs(title = "Resturant Insepction Violations",x="Borough",y="Violations")
```

# Data 2 - Police Killings Since 2015

## Introduction
This dataset is of all Police killings from January 1, 2015 to September 27, 2020. It was collected by The Washington Post and can be found on their [GitHub](https://github.com/washingtonpost/data-police-shootings). 

### Get Dataset
First we pull the most recent csv file and create a dataframe. 
```{r}
#get data file
site2 <- getURL("https://raw.githubusercontent.com/ltcancel/Project2_Data607_F20/master/fatal-police-shootings-data.csv")

df2 <- read.csv(text = site2,sep = ",",header = TRUE)
head(df2)

```
### Tidy the data/Analysis
Q: Which states had the most deadly shootings from 2015 - 2020

I tidied the data by splitting the date into 3 columns so I can do an analysis by year. Since this is such a large dataset, I took a subset of data that only includes the 5 states with the highest number of killings for each year. 
```{r}
#Split date into year, month, and day columns so we can makes graphs based on year
df2<- df2 %>%
  separate(date,sep="-",into = c("year","month","day"),remove = FALSE)
head(df2)

#top 5 states with the highest number of deaths per year
top5 <- df2 %>% group_by(year) %>% count(state) %>% top_n(5)
head(top5)
```

In the bar graph below, we can see that some states are on the top of the list for all 6 years. California has the greatest number of police killings each year, followed by Texas, and then Florida. 2020 is the only year where Florida has slightly more killings than Texas. Arizona has dropped from the top 5 in 2019. 
```{r}
#Barplot using the top 5 states with highest number of deaths per year data
ggplot(top5) +
  geom_bar(mapping = aes(x = state, y = n, fill = state), stat = 'identity') +
  facet_wrap(year~.) +
  labs(title = "Top 5 States with Deadly Shootings")
```

Q: What is the average number of police killings by city?
```{r}
#get average shootings per year
yearCount <- df2 %>% group_by(year) %>% count(state)


stateAvg <- yearCount %>% 
  group_by(state) %>%
  summarise(Average = mean(n)) %>%
  arrange(desc(Average))

head(stateAvg)
```

This graph is a bit crowded with bars, but we can clearly see that California and Texas has the greatest average. 
```{r}
ggplot(stateAvg) +
  geom_bar(mapping = aes(x = state, y = Average, fill = state), stat = 'identity') +
  coord_flip() +
  scale_y_continuous(breaks = seq(0,150,25))
```

Q: How many of the victims showed signs of mental illness

Each year we can see a large gap in the number of killings where the victim showed signs of mental illness versus those who did not. Most were reported to not have shown signs of mental illness. 
```{r}
ggplot(df2) +
  geom_bar(mapping = aes(x=year,fill=signs_of_mental_illness), position = "dodge")
```

Since the count looks very similar year over year, I created a subset of states with the highest total number of killings. There isn’t much of a significant difference. Each of the states has a gap between victims who showed signs of mental illness versus those who did not. 
```{r}
df2Filter <- df2 %>%
  filter(state %in% c("AZ","CA","CO","FL","TX"))

ggplot(df2Filter) +
  geom_bar(mapping = aes(x=year, fill=signs_of_mental_illness), position = "dodge") +
  facet_wrap(~state) +
  coord_flip()
```

 

